Hydrodynamics of a Semipermeable Inextensible Membrane Under Flow and Confinement

### 2021-05-19 vesicle_code/docs/permeableVes/semipermeable_vesicle.tex
yny

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Role of Tension and Bending on Semipermeability}
\textcolor{magenta}{Since $\jump{p}= \ff_\mathrm{mem} \cdot
\nn$~\cite{poz1992}, our model assumes that the permeating flux is
proportional to the mechanical pressure jump across the membrane. We
decompose the normal component of the membrane force into its
contributions due to tension and bending as
\begin{align}
  \ff_{\mathrm{ten}} \cdot \nn = -\kappa \Lambda 
    \quad \text{ and } \quad
  \ff_{\mathrm{ben}} \cdot \nn = \kappa_{ss} + \frac{1}{2}\kappa^3.
\end{align}
Then, the total flux through the membrane due to the tension and bending
forces are
\begin{align}
  \dot{A}_{\mathrm{ten}}(t) = -\beta \int_{\gamma} \kappa \Lambda\, ds  
  \quad \text{ and } \quad
  \dot{A}_{\mathrm{ben}}(t) = \frac{\beta}{2} \int_{\gamma} \kappa^3\, ds,
  \label{eqn:areaROCdecomposed}
\end{align}
respectively. Note that the sum of $\dot{A}_{\mathrm{ten}}(t)$ and
$\dot{A}_{\mathrm{ben}}(t)$ results in equation~\eqref{eqn:areaROC}.
After multiplying~\eqref{eqn:areaROCdecomposed} by $4\pi/L^2$, where $L$
is the constant  vesicle length, the rate of change of the reduced area
due to the tension and bending forces are formed. We investigate the
effect of the tension and bending forces on the semipermeability for two
quiescent examples (Figures~\ref{fig:unconfinedRAdecomp}(a) and (b)),
one planar shear example (Figure~\ref{fig:unconfinedRAdecomp}(c)), and
one Poiseuille example (Figure~\ref{fig:unconfinedRAdecomp}(d)). In the
quiescent flow examples, we see that the tension and bending forces
result in similar amounts of semipermeability, and for the geometry with
a larger initial curvature, the reduced area changes much quicker and
the vesicle deflates at early times before inflating to a circular
shape. In the planar shear flow and Poiseuille flow examples, the
tension and bending forces result in similar amounts of
semipermeability.}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/unconfinedRAdecomp.tikz}
  \else
  \includegraphics{figures/unconfinedRAdecomp.pdf}
  \fi
  \caption{\label{fig:unconfinedRAdecomp}\textcolor{magenta}{The rate of
  change of the reduced area caused by the tension and bending forces.
  The dashed line is the sum of these two contributions. (a) The
  quiescent example in Figure~\ref{fig:relaxationComposite} with
  $\alpha(0) = 0.66$. (b) The quiescent example in
  Figure~\ref{fig:relaxationComposite} with $\alpha(0) = 0.38$. (c) The
  planar shear example in Figure~\ref{fig:shearComposite} with $\alpha(0) =
  0.45$ and $Ca_\mathrm{E} = 10^0$. (d) The Poiseuille example in
  Figure~\ref{fig:parabolicComposite} with $\alpha(0) = 0.60$ and
  maximum velocity $U = 600~\mu$m/s.}}
\end{figure*}

@@

\textcolor{magenta}{Using equation~\eqref{eqn:areaROCdecomposed}, we
investigate the effects of the tension and bending forces on the water
flux through the membrane. In Figure~\ref{fig:stenosisRAdecomp}, we
compute the rate of change of the reduced area due to the tension force
(blue), the bending force (red), and the total membrane force (black).
In both the short and long closely-fit channels, the tension force is
responsible for nearly all the semipermeability. This is in contrast to
the unconfined examples where the tension and bending forces were
each responsible for a comparable amount of semipermeability.}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/stenosisRAdecomp.tikz}
  \else
  \includegraphics{figures/stenosisRAdecomp.pdf}
  \fi
  \caption{\label{fig:stenosisRAdecomp} \textcolor{magenta}{The rate of
  change of the reduced area caused by the tension and bending forces.
  The dashed line is the sum of these two contributions. (a) The short
  stenosis geometry in Figure~\ref{fig:stenosisComposite}(a). (b) The
  long stenosis geometry in Figure~\ref{fig:stenosisComposite}(b).}}
\end{figure*}

@@

\textcolor{magenta}{As was done for the closely-fit channel, we use
equation~\eqref{eqn:areaROCdecomposed} to investigate the effects of the
tension and bending forces on semipermeability.
Figure~\ref{fig:contractingRAdecomp} shows the rate of change of the
reduced area due to the tension force (blue), the bending force (red),
and the total membrane force (black). Again, the tension is responsible
for the majority of the semipermeability. However, when compared to the
closely-fit channel example, the bending does have more of a
contribution because of the lower maximum imposed velocity. While we
only decompose the membrane force into tension and bending for the two
vesicles that successfully pass through the contraction, we also plot
the total membrane force for the two vesicles that do not pass through
the contraction (green). The vesicle does deflate with a smaller
semipermeability coefficient and a smaller maximum imposed velocity, but
the rate of efflux is insufficient for successful passage of the vesicle
through the contraction.}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/contractingRAdecomp.tikz}
  \else
  \includegraphics{figures/contractingRAdecomp.pdf}
  \fi
  \caption{\label{fig:contractingRAdecomp} \textcolor{magenta}{The rate
  of change of the reduced area caused by the tension and bending
  forces. The dashed line is the sum of these two contributions. Only
  simulations where the vesicle passes through the contraction are
  decomposed into the contributions due to tension and bending. (a) The
  low semipermeability coefficient in
  Figure~\ref{fig:contractingComposite}(b). The green line corresponds
  to the simulation in Figure~\ref{fig:contractingComposite}(a). (b) The
  high semipermeability coefficient in
  Figure~\ref{fig:contractingComposite}(e). The green line corresponds
  to the simulation in Figure~\ref{fig:contractingComposite}(d).}}
\end{figure*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contribution from Membrane Tension and Bending to the Permeating Flux}
\textcolor{magenta}{In the absence of viscosity contrast between interior and exterior fluids, the pressure jump $\jump{p}= \ff_\mathrm{mem} \cdot
\nn$~\cite{poz1992}. Thus in our model, the permeating flux is
proportional to the normal component of membrane forces $\ff_\mathrm{mem}\cdot\nn$. We
decompose $\ff_\mathrm{mem}\cdot\nn$ into its
contributions due to tension and bending as
\begin{align}
  \ff_{\mathrm{ten}} \cdot \nn = -\kappa \Lambda 
    \quad \text{ and } \quad
  \ff_{\mathrm{ben}} \cdot \nn = \kappa_{ss} + \frac{1}{2}\kappa^3.
\end{align}
Then, the total flux through the membrane due to the tension and bending
forces are
\begin{align}
  \dot{A}_{\mathrm{ten}}(t) = -\beta \int_{\gamma} \kappa \Lambda\, ds  
  \quad \text{ and } \quad
  \dot{A}_{\mathrm{ben}}(t) = \frac{\beta}{2} \int_{\gamma} \kappa^3\, ds,
  \label{eqn:areaROCdecomposed}
\end{align}
respectively. Note that the sum of $\dot{A}_{\mathrm{ten}}(t)$ and
$\dot{A}_{\mathrm{ben}}(t)$ results in equation~\eqref{eqn:areaROC}.
Multiplying~\eqref{eqn:areaROCdecomposed} by $4\pi/L^2$, where $L$
is the constant  vesicle length, we obtain the rate of change of the reduced area
due to the tension and bending forces. 
Figures~\ref{fig:unconfinedRAdecomp} summarize the 
effect of the tension and bending forces on the semipermeability for two
quiescent examples (Figures~\ref{fig:unconfinedRAdecomp}(a) and (b)),
one planar shear example (Figure~\ref{fig:unconfinedRAdecomp}(c)), and
one Poiseuille example (Figure~\ref{fig:unconfinedRAdecomp}(d)). In the
quiescent flow examples, we see that the tension and bending forces
result in similar amounts of semipermeability, and for the geometry with
a larger initial curvature, the reduced area changes much quicker and
the vesicle deflates at early times before inflating to a circular
shape. In the planar shear flow and Poiseuille flow examples, the
tension and bending forces result in similar amounts of
semipermeability.}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/unconfinedRAdecomp.tikz}
  \else
  \includegraphics{figures/unconfinedRAdecomp.pdf}
  \fi
  \caption{\label{fig:unconfinedRAdecomp}\textcolor{magenta}{The rate of
  change of the reduced area decomposed into contributions from membrane tension (blue curves) and membrane bending (red curves).
  The dashed line is the sum of these two contributions. (a) The
  quiescent example in Figure~\ref{fig:relaxationComposite} with
  $\alpha(0) = 0.66$. (b) The quiescent example in
  Figure~\ref{fig:relaxationComposite} with $\alpha(0) = 0.38$. (c) The
  planar shear example in Figure~\ref{fig:shearComposite} with $\alpha(0) =
  0.45$ and $Ca_\mathrm{E} = 10^0$. (d) The Poiseuille example in
  Figure~\ref{fig:parabolicComposite} with $\alpha(0) = 0.60$ and
  maximum velocity $U = 600~\mu$m/s.}}
\end{figure*}

@@

\textcolor{magenta}{
Figure~\ref{fig:stenosisRAdecomp} show
the rate of change of the reduced area due to the tension force
(blue), the bending force (red), and the total membrane force (black).
%Using equation~\eqref{eqn:areaROCdecomposed}, we
%investigate the effects of the tension and bending forces on the water
%flux through the membrane. 
In both short and long closely-fit channels, the membrane tension is
responsible for nearly all the permeating water flux. This is in contrast to
the unconfined examples where the tension and bending forces were
each responsible for a comparable amount of permeating water flux.}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/stenosisRAdecomp.tikz}
  \else
  \includegraphics{figures/stenosisRAdecomp.pdf}
  \fi
  \caption{\label{fig:stenosisRAdecomp} \textcolor{magenta}{The rate of
  change of the reduced area due to the membrane tension (blue curves) and membrane bending (red curves).
  The dashed line is the sum of these two contributions. (a) The short
  stenosis geometry in Figure~\ref{fig:stenosisComposite}(a). (b) The
  long stenosis geometry in Figure~\ref{fig:stenosisComposite}(b).}}
\end{figure*}

@@

\textcolor{magenta}{
%As was done for the closely-fit channel, we use
%equation~\eqref{eqn:areaROCdecomposed} to investigate the effects of the
%tension and bending forces on semipermeability.
Figure~\ref{fig:contractingRAdecomp} shows the rate of change of the
reduced area due to membrane tension (blue curves), membrane bending (red curves),
and the total membrane force (dashed black curves). Again, the tension is responsible
for the majority of the permeating flux. However, when compared to the
closely-fit channel example, the bending does have more of a
contribution because of the lower maximum imposed velocity. While we
only decompose the membrane force into tension and bending for the two
vesicles that successfully pass through the contraction, we also plot
the total membrane force for the two vesicles that do not pass through
the contraction (dash-dotted green curves). The vesicle does deflate with a smaller
semipermeability coefficient and a smaller maximum imposed velocity, but
the rate of efflux is insufficient for successful passage of the vesicle
through the contraction.}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/contractingRAdecomp.tikz}
  \else
  \includegraphics{figures/contractingRAdecomp.pdf}
  \fi
  \caption{\label{fig:contractingRAdecomp} \textcolor{magenta}{The rate
  of change of the reduced area due to  membrane tension (blue curves) and membrane bending
  (red curves). The dashed black  line is the sum of these two contributions. Only
  simulations where the vesicle passes through the contraction are
  decomposed into the contributions due to tension and bending. (a) The
  low semipermeability coefficient in
  Figure~\ref{fig:contractingComposite}(b). The dash-dotted green curve corresponds
  to the simulation in Figure~\ref{fig:contractingComposite}(a). (b) The
  high semipermeability coefficient in
  Figure~\ref{fig:contractingComposite}(e). The dash-dotted green curve corresponds
  to the simulation in Figure~\ref{fig:contractingComposite}(d).}}
\end{figure*}
